---
title: "11. SVM"
author: "Al Xin"
date: "1/5/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = F)
library(pacman)
p_load(
  caret,
  data.table,
  doParallel, 
  dplyr, 
  ggplot2, 
  parallel, 
  qsarr,
  stringr, 
  tidyr
)
theme_set(theme_bw())
knitr::opts_chunk$set(echo = TRUE, eval = F)
```

## Overview

Models built on all predictor variables will be saved in `model/svm_polynomial/all_var/`. 
Models trained on variables selected from RFE in 09 will be saved in `model/svm_polynomial/rfe_var`. 

```{r directories}
desc_src <- c("cdk", "mordred", "ochem", "padel")

if (!dir.exists("model")) dir.create("model")
if (!dir.exists("model/svm_polynomial")) {
  dir.create("model/svm_polynomial")
  dir.create("model/svm_polynomial/all_var")
  dir.create("model/svm_polynomial/rfe_var")
  dir_create_result <- sapply(paste0("model/svm_polynomial/all_var/", desc_src), dir.create)
  dir_create_result <- sapply(paste0("model/svm_polynomial/rfe_var/", desc_src), dir.create)
}
```

### General outline for model building

## All variables

For creating models with all the variables, the descriptors are read as separate 
lists. This makes it clearer that different possible tuning parameters are used
based on the number of descriptors available. 

```{r cdk_all}
# Read preprocessed descriptors
# Remove guest names for training
cdk_list <- read_desc_list(
  list.files("preprocess/cdk", full.names = T), 
  quiet = T
)

padel_list <- read_desc_list(
  list.files("preprocess/padel", full.names = T), 
  quiet = T
)

mordred_list <- read_desc_list(
  list.files("preprocess/mordred", full.names = T), 
  quiet = T
)

ochem_list <- read_desc_list(
  list.files("preprocess/ochem", full.names = T), 
  quiet = T
) 

data_name <- names(cdk_list)
```

## Polynomial SVM

Because the SVM kernels all have slightly different parameters, `qsarr` handles each kernel through a different `tune` call. It is more straightforward to feed the `tune` function the parameters relevant to each kernel rather than check dependencies on the kernel type. 

This may change in later versions of `qsarr` but is true for v0.1.0.

An SVM with a polynomial kernel draws a linear boundary to explain the data. 

### All Variables

```{r, eval = F}
num_core <- max(1, detectCores() - 1)
cl <- makeCluster(num_core)
registerDoParallel(cl)
cl_lib <- clusterEvalQ(cl, library("qsarr"))

clusterSetRNGStream(cl = cl, 20210807)
cdk_all <- parLapply(
  cl = cl,
  cdk_list,
  tune,
  method = "svm_polynomial",
  resp = "dG",
  ignore_col = "guest",
  nfold = 10,
  nrep = 10,
  degree = c(1:5),
  coef0 = c(-10, -1, 0, 1, 10),
  cost = c(0.5, 1, 3, 10, 20)
)

cdk_all_save <- lapply(
  data_name,
  function(x) {
    saveRDS(cdk_all[[x]], paste0("model/svm_polynomial/all_var/cdk/", x, ".RDS"))
  }
)

clusterSetRNGStream(cl = cl, 20210807)
padel_all <- parLapply(
  cl = cl,
  padel_list, 
  tune, 
  method = "svm_polynomial", 
  resp = "dG",
  ignore_col = "guest",
  nfold = 10,
  nrep = 10,
  degree = c(1:5),
  coef0 = c(-10, -1, 0, 1, 10),
  cost = c(0.5, 1, 3, 10, 20)
)

padel_all_save <- lapply(
  data_name, 
  function(x) {
    saveRDS(padel_all[[x]], paste0("model/svm_polynomial/all_var/padel/", x, ".RDS"))
  }
)

clusterSetRNGStream(cl = cl, 20210807)
mordred_all <- parLapply(
  cl = cl,
  mordred_list, 
  tune, 
  method = "svm_polynomial", 
  resp = "dG",
  ignore_col = "guest",
  nfold = 10,
  nrep = 10,
  degree = c(1:5),
  coef0 = c(-10, -1, 0, 1, 10),
  cost = c(0.5, 1, 3, 10, 20)
)

mordred_all_save <- lapply(
  data_name, 
  function(x) {
    saveRDS(mordred_all[[x]], paste0("model/svm_polynomial/all_var/mordred/", x, ".RDS"))
  }
)

clusterSetRNGStream(cl = cl, 20210807)
ochem_all <- parLapply(
  cl = cl,
  ochem_list, 
  tune, 
  method = "svm_polynomial", 
  resp = "dG",
  ignore_col = "guest",
  nfold = 10,
  nrep = 10,
  degree = c(1:5),
  coef0 = c(-10, -1, 0, 1, 10),
  cost = c(0.5, 1, 3, 10, 20)
)

ochem_all_save <- lapply(
  data_name, 
  function(x) {
    saveRDS(ochem_all[[x]], paste0("model/svm_polynomial/all_var/ochem/", x, ".RDS"))
  }
)
stopCluster(cl = cl)
```

#### RFE

```{r rfe}
# Read preprocessed descriptors
# Remove guest names for training
cdk_list_rfe <- read_desc_list(
  list.files("rfe/df/cdk", full.names = T), 
  quiet = T
)

padel_list_rfe <- read_desc_list(
  list.files("rfe/df/padel", full.names = T), 
  quiet = T
)

mordred_list_rfe <- read_desc_list(
  list.files("rfe/df/mordred", full.names = T), 
  quiet = T
)

ochem_list_rfe <- read_desc_list(
  list.files("rfe/df/ochem", full.names = T), 
  quiet = T
) 

data_name <- names(cdk_list_rfe)
```

```{r}
# Parallel processing setup
num_core <- max(1, detectCores() - 1)
cl <- makeCluster(num_core)
registerDoParallel(cl)
cl_lib <- clusterEvalQ(cl, library("qsarr"))

clusterSetRNGStream(cl = cl, 20210722)
cdk_rfe <- parLapply(
  cl = cl,
  cdk_list_rfe,
  tune, 
  method = "svm_polynomial", 
  resp = "dG",
  ignore_col = "guest",
  nfold = 10,
  nrep = 10,
  degree = c(1:5),
  coef0 = c(-10, -1, 0, 1, 10),
  cost = c(0.5, 1, 3, 10, 20)
)

cdk_rfe_save <- lapply(
  data_name, 
  function(x) {
    saveRDS(cdk_rfe[[x]], paste0("model/svm_polynomial/rfe_var/cdk/", x, ".RDS"))
  }
)

clusterSetRNGStream(cl = cl, 20210722)
padel_rfe <- parLapply(
  cl = cl,
  padel_list_rfe,
  tune, 
  method = "svm_polynomial", 
  resp = "dG",
  ignore_col = "guest",
  nfold = 10,
  nrep = 10,
  degree = c(1:5),
  coef0 = c(-10, -1, 0, 1, 10),
  cost = c(0.5, 1, 3, 10, 20)
)

padel_rfe_save <- lapply(
  data_name, 
  function(x) {
    saveRDS(padel_rfe[[x]], paste0("model/svm_polynomial/rfe_var/padel/", x, ".RDS"))
  }
)

clusterSetRNGStream(cl = cl, 20210722)
mordred_rfe <- parLapply(
  cl = cl,
  mordred_list_rfe,
  tune, 
  method = "svm_polynomial", 
  resp = "dG",
  ignore_col = "guest",
  nfold = 10,
  nrep = 10,
  degree = c(1:5),
  coef0 = c(-10, -1, 0, 1, 10),
  cost = c(0.5, 1, 3, 10, 20)
)

mordred_rfe_save <- lapply(
  data_name, 
  function(x) {
    saveRDS(mordred_rfe[[x]], paste0("model/svm_polynomial/rfe_var/mordred/", x, ".RDS"))
  }
)

clusterSetRNGStream(cl = cl, 20210722)
ochem_rfe <- parLapply(
  cl = cl,
  ochem_list_rfe,
  tune, 
  method = "svm_polynomial", 
  resp = "dG",
  ignore_col = "guest",
  nfold = 10,
  nrep = 10,
  degree = c(1:5),
  coef0 = c(-10, -1, 0, 1, 10),
  cost = c(0.5, 1, 3, 10, 20)
)

ochem_rfe_save <- lapply(
  data_name, 
  function(x) {
    saveRDS(ochem_rfe[[x]], paste0("model/svm_polynomial/rfe_var/ochem/", x, ".RDS"))
  }
)
stopCluster(cl = cl)
```

### Polynomial SVMs

```{r}
if (!dir.exists("model/svm_polynomial")) {
  dir.create("model/svm_polynomial/all_var", recursive = T)
  dir.create("model/svm_polynomial/rfe_var")
  dir_create_result <- sapply(paste0("model/svm_polynomial/all_var/", desc_src), dir.create)
  dir_create_result <- sapply(paste0("model/svm_polynomial/rfe_var/", desc_src), dir.create)
}
```

#### All Variables

```{r}
# Rerun this each kernel because this will likely take multiple sittings
num_core <- max(1, detectCores() - 1)
cl <- makeCluster(num_core)
registerDoParallel(cl)
cl_lib <- clusterEvalQ(cl, library("qsarr"))

clusterSetRNGStream(cl = cl, 20210722)

cdk_all <- parLapply(
  cl = cl,
  cdk_list, 
  tune, 
  method = "svm_polynomial", 
  resp = "dG",
  ignore_col = "guest",
  nfold = 10,
  nrep = 10,
  degree = c(1:5),
  coef0 = c(-10, -1, 0, 1, 10),
  cost = c(0.5, 1, 3, 10, 20), 
  epsilon = c(0.01, 0.1, 0.5, 0.9), 
  gamma = c(0.001, 0.01, 0.1, 0.25, 0.5)
)

cdk_all_save <- lapply(
  data_name, 
  function(x) {
    saveRDS(cdk_all[[x]], paste0("model/svm_polynomial/all_var/cdk/", x, ".RDS"))
  }
)

clusterSetRNGStream(cl = cl, 20210722)
padel_all <- parLapply(
  cl = cl,
  padel_list, 
  tune, 
  method = "svm_polynomial", 
  resp = "dG",
  ignore_col = "guest",
  nfold = 10,
  nrep = 10,
  degree = c(1:5),
  coef0 = c(-10, -1, 0, 1, 10),
  cost = c(0.5, 1, 3, 10, 20), 
  epsilon = c(0.01, 0.1, 0.5, 0.9), 
  gamma = c(0.001, 0.01, 0.1, 0.25, 0.5)
)

padel_all_save <- lapply(
  data_name, 
  function(x) {
    saveRDS(padel_all[[x]], paste0("model/svm_polynomial/all_var/padel/", x, ".RDS"))
  }
)

clusterSetRNGStream(cl = cl, 20210722)
mordred_all <- parLapply(
  cl = cl,
  mordred_list, 
  tune, 
  method = "svm_polynomial", 
  resp = "dG",
  ignore_col = "guest",
  nfold = 10,
  nrep = 10,
  degree = c(1:5),
  coef0 = c(-10, -1, 0, 1, 10),
  cost = c(0.5, 1, 3, 10, 20), 
  epsilon = c(0.01, 0.1, 0.5, 0.9), 
  gamma = c(0.001, 0.01, 0.1, 0.25, 0.5)
)

mordred_all_save <- lapply(
  data_name, 
  function(x) {
    saveRDS(mordred_all[[x]], paste0("model/svm_polynomial/all_var/mordred/", x, ".RDS"))
  }
)

clusterSetRNGStream(cl = cl, 20210722)
ochem_all <- parLapply(
  cl = cl,
  ochem_list, 
  tune, 
  method = "svm_polynomial", 
  resp = "dG",
  ignore_col = "guest",
  nfold = 10,
  nrep = 10,
  degree = c(1:5),
  coef0 = c(-10, -1, 0, 1, 10),
  cost = c(0.5, 1, 3, 10, 20), 
  epsilon = c(0.01, 0.1, 0.5, 0.9), 
  gamma = c(0.001, 0.01, 0.1, 0.25, 0.5)
)

ochem_all_save <- lapply(
  data_name, 
  function(x) {
    saveRDS(ochem_all[[x]], paste0("model/svm_polynomial/all_var/ochem/", x, ".RDS"))
  }
)
stopCluster(cl = cl)
```

#### RFE

```{r}
num_core <- max(1, detectCores() - 1)
cl <- makeCluster(num_core)
registerDoParallel(cl)
cl_lib <- clusterEvalQ(cl, library("qsarr"))

clusterSetRNGStream(cl = cl, 20210722)

cdk_rfe <- parLapply(
  cl = cl,
  cdk_list_rfe, 
  tune, 
  method = "svm_polynomial", 
  resp = "dG",
  ignore_col = "guest",
  nfold = 10,
  nrep = 10,
  degree = c(1:5),
  coef0 = c(-10, -1, 0, 1, 10),
  cost = c(0.5, 1, 3, 10, 20), 
  epsilon = c(0.01, 0.1, 0.5, 0.9), 
  gamma = c(0.001, 0.01, 0.1, 0.25, 0.5)
)

cdk_rfe_save <- lapply(
  data_name, 
  function(x) {
    saveRDS(cdk_rfe[[x]], paste0("model/svm_polynomial/rfe_var/cdk/", x, ".RDS"))
  }
)

clusterSetRNGStream(cl = cl, 20210722)
padel_rfe <- parLapply(
  cl = cl,
  padel_list_rfe, 
  tune, 
  method = "svm_polynomial", 
  resp = "dG",
  ignore_col = "guest",
  nfold = 10,
  nrep = 10,
  degree = c(1:5),
  coef0 = c(-10, -1, 0, 1, 10),
  cost = c(0.5, 1, 3, 10, 20), 
  epsilon = c(0.01, 0.1, 0.5, 0.9), 
  gamma = c(0.001, 0.01, 0.1, 0.25, 0.5)
)

padel_rfe_save <- lapply(
  data_name, 
  function(x) {
    saveRDS(padel_rfe[[x]], paste0("model/svm_polynomial/rfe_var/padel/", x, ".RDS"))
  }
)

clusterSetRNGStream(cl = cl, 20210722)
mordred_rfe <- parLapply(
  cl = cl,
  mordred_list_rfe, 
  tune, 
  method = "svm_polynomial", 
  resp = "dG",
  ignore_col = "guest",
  nfold = 10,
  nrep = 10,
  degree = c(1:5),
  coef0 = c(-10, -1, 0, 1, 10),
  cost = c(0.5, 1, 3, 10, 20), 
  epsilon = c(0.01, 0.1, 0.5, 0.9), 
  gamma = c(0.001, 0.01, 0.1, 0.25, 0.5)
)

mordred_rfe_save <- lapply(
  data_name, 
  function(x) {
    saveRDS(mordred_rfe[[x]], paste0("model/svm_polynomial/rfe_var/mordred/", x, ".RDS"))
  }
)

clusterSetRNGStream(cl = cl, 20210722)
ochem_rfe <- parLapply(
  cl = cl,
  ochem_list_rfe, 
  tune, 
  method = "svm_polynomial", 
  resp = "dG",
  ignore_col = "guest",
  nfold = 10,
  nrep = 10,
  degree = c(1:5),
  coef0 = c(-10, -1, 0, 1, 10),
  cost = c(0.5, 1, 3, 10, 20), 
  epsilon = c(0.01, 0.1, 0.5, 0.9), 
  gamma = c(0.001, 0.01, 0.1, 0.25, 0.5)
)

ochem_rfe_save <- lapply(
  data_name, 
  function(x) {
    saveRDS(ochem_rfe[[x]], paste0("model/svm_polynomial/rfe_var/ochem/", x, ".RDS"))
  }
)
stopCluster(cl = cl)
```

### Radial basis function

#### All Variables

```{r}

```

#### RFE

```{r}

```

## External validation

```{r}
desc_name <- c("cdk", "mordred", "ochem", "padel")
data_name <- str_remove(list.files("desc/cdk"), "\\.RDS$")
```

First, the data generated from Step 07 is read into a list.

```{r}
ev_list <- lapply(
  paste0("extval_pp/", desc_name), 
  function(x) {
    lapply(
      list.files(x, full.names = T), 
      readRDS
    ) %>%
      setNames(nm = data_name)
  }
) %>%
  setNames(nm = desc_name)
```

External validation follows the same steps as described in `10-rf.Rmd`. 

### Linear SVMs

#### All Variables

The results on external validation for linear SVMs can be predicted. Performance on all molecules and performance on the applicability domain can be contrasted.

```{r}
svm_polynomial_allvar <- lapply(
  paste0("model/svm_polynomial/all_var/", desc_name), 
  function(x) {
    lapply(
      list.files(x, full.names = T), 
      readRDS
    ) %>%
      setNames(nm = data_name)
  }
) %>%
  setNames(nm = desc_name)
```

```{r}
ev_allvar <- lapply(
  desc_name, 
  function(x) {
    tune_sublist <- svm_polynomial_allvar[[x]]
    ev_sublist <- ev_list[[x]]
    lapply(
      data_name, 
      function(y) {
        pred <- predict(tune_sublist[[y]], ev_sublist[[y]])
        data.frame(
          guest = ev_sublist[[y]]$guest, 
          in_ad = ev_sublist[[y]]$in_ad,
          obs = ev_sublist[[y]]$dG, 
          pred = pred
        )
      }
    ) %>%
      setNames(data_name)
  }
) %>% 
  setNames(desc_name)

# All results, regardless of applicability domain
ev_allvar_result <- lapply(
  ev_allvar, 
  function(x) {
    lapply(
      x, 
      defaultSummary
    ) %>%
      setNames(data_name)
  }
) %>%
  setNames(desc_name)

# Accuracy only on applicability domain
ev_allvar_result_ad <- lapply(
  ev_allvar, 
  function(x) {
    lapply(
      x, 
      function(y) {
        y <- y[y$in_ad, ]
        defaultSummary(y)
      }
    ) %>%
      setNames(data_name)
  }
) %>% 
  setNames(desc_name)
```

#### RFE

The same procedure for prediction is carried out with models only built on variables selected with RFE. 

```{r}
svm_polynomial_rfe <- lapply(
  paste0("model/svm_polynomial/rfe_var/", desc_name), 
  function(x) {
    lapply(
      list.files(x, full.names = T), 
      readRDS
    ) %>%
      setNames(nm = data_name)
  }
) %>%
  setNames(nm = desc_name)

ev_rfe <- lapply(
  desc_name, 
  function(x) {
    tune_sublist <- svm_polynomial_rfe[[x]]
    ev_sublist <- ev_list[[x]]
    lapply(
      data_name, 
      function(y) {
        pred <- predict(tune_sublist[[y]], ev_sublist[[y]])
        data.frame(
          guest = ev_sublist[[y]]$guest, 
          in_ad = ev_sublist[[y]]$in_ad,
          obs = ev_sublist[[y]]$dG, 
          pred = pred
        )
      }
    ) %>%
      setNames(data_name)
  }
) %>% 
  setNames(desc_name)

# All results, regardless of applicability domain
ev_rfe_result <- lapply(
  ev_rfe, 
  function(x) {
    lapply(
      x, 
      defaultSummary
    ) %>%
      setNames(data_name)
  }
) %>%
  setNames(desc_name)

# Accuracy only on applicability domain
ev_rfe_result_ad <- lapply(
  ev_rfe, 
  function(x) {
    lapply(
      x, 
      function(y) {
        y <- y[y$in_ad, ]
        defaultSummary(y)
      }
    ) %>%
      setNames(data_name)
  }
) %>% 
  setNames(desc_name)
```

## Polynomial SVM



## Results

### Model-building process

The results are first read from the lists generated from tuning and saved in the
directory `result/trn/`.

```{r}
svm_polynomial_allvar_df <- rbindlist(
  lapply(
    desc_name, 
    function(x) {
      desc_list <- svm_polynomial_allvar[[x]]
      rbindlist(
        lapply(
          data_name, 
          function(y) {
            mutate(desc_list[[y]]$result_summary, data_src = y)
          }
        )
      ) %>%
        mutate(desc_src = x)
    }
  )
) 

svm_polynomial_rfe_df <- rbindlist(
  lapply(
    desc_name, 
    function(x) {
      desc_list <- svm_polynomial_rfe[[x]]
      rbindlist(
        lapply(
          data_name, 
          function(y) {
            mutate(desc_list[[y]]$result_summary, data_src = y)
          }
        )
      ) %>%
        mutate(desc_src = x)
    }
  )
) 

saveRDS(svm_polynomial_allvar_df, "result/trn/all_var/rf.RDS")
saveRDS(svm_polynomial_rfe_df, "result/trn/rfe_var/rf.RDS")
```

```{r, eval = T, echo = F}
svm_polynomial_allvar_df <- readRDS("result/trn/all_var/rf.RDS")
svm_polynomial_rfe_df <- readRDS("result/trn/rfe_var/rf.RDS")

ggplot(svm_polynomial_allvar_df, aes(x = data_src, y = fold_avg, fill = desc_src)) + 
  geom_col(position = "dodge") + 
  facet_grid(summary_stat ~ ., scales = "free") + 
  labs(
    x = "Data source", 
    y = "Average value over repeated 10-fold CV", 
    title = "Results for random forest models on all variables", 
    fill = "Chemical descriptor"
  )

ggplot(svm_polynomial_rfe_df, aes(x = data_src, y = fold_avg, fill = desc_src)) + 
  geom_col(position = "dodge") + 
  facet_grid(summary_stat ~ ., scales = "free") + 
  labs(
    x = "Data source", 
    y = "Average value over repeated 10-fold CV", 
    title = "Results for random forest models on RFE-selected variables", 
    fill = "Chemical descriptor"
  )
```

### External validation 

```{r}
ev_allvar_df <- data.table::rbindlist(
  lapply(
    unlist(ev_allvar_result, recursive = F), 
    function(x) {
      data.frame(as.list(x))
    }
  )
) %>%
  data.frame(
    desc_src = unlist(lapply(desc_name, rep, length(data_name))), 
    data_src = rep(data_name, length(desc_name))
  ) %>%
  pivot_longer(RMSE:MAE, names_to = "summary_stat")

ev_allvar_ad_df <- data.table::rbindlist(
  lapply(
    unlist(ev_allvar_result_ad, recursive = F), 
    function(x) {
      data.frame(as.list(x))
    }
  )
) %>%
  data.frame(
    desc_src = unlist(lapply(desc_name, rep, length(data_name))), 
    data_src = rep(data_name, length(desc_name))
  ) %>%
  pivot_longer(RMSE:MAE, names_to = "summary_stat")

ggplot(ev_allvar_df, aes(x = data_src, y = value, fill = desc_src)) + 
  geom_col(position = "dodge") + 
  facet_grid(summary_stat ~ ., scales = "free")

ggplot(ev_allvar_ad_df, aes(x = data_src, y = value, fill = desc_src)) + 
  geom_col(position = "dodge") + 
  facet_grid(summary_stat ~ ., scales = "free")
# svm_polynomial_tst <- svm_polynomial_allvar[[1]]
# ev_tst <- ev_list[[1]]
# pred_tst <- lapply(
#   names(svm_polynomial_tst), 
#   function(x) {
#     # svm_polynomial_model <- svm_polynomial_tst[[x]]$model
#     # predict(svm_polynomial_model, ev_tst[[x]])
#     predict(svm_polynomial_tst[[x]], ev_tst[[x]])
#   }
# )
```

```{r}
print.tune <- function(tune_obj) {
  cat(
    "Model type: ", class(tune_obj$model), "\n\n",
    "Tuned parameters: ", "\n",
    sapply(
      names(tune_obj$param),
      function(x) {
        paste0("\t", x, ": ", tune_obj$param[[x]], "\n")
      }
    ), "\n",
    "Number of predictors: ", length(tune_obj$pred_name), "\n",
    "Predictors (first 10): ",
    tune_obj$pred_name[1:(min(10, length(tune_obj$pred_name)))],
    "\n", "\n",
    "Model performance: ", "\n"
  )
  print(tune_obj$result_summary)
}

predict.tune <- function(tune_obj, ...) {
  # if (length(class(tune_obj$model)) > 1) {
  #   class(tune_obj$model) <-
  #     class(tune_obj$model)[length(class(tune_obj$model))]
  # }
  # best_model <- tune_obj$model
  # UseMethod("predict", best_model)
  predict(tune_obj$model, ...)
}
```

